---
title: "Joining your own data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Joining your own data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE,
  fig.width = 7, fig.height = 4, fig.align = "center", dpi = 96
)
library(worlddatajoin)
library(ggplot2)
library(dplyr)
```

The headline use case: *I have a frame keyed on messy country names — get it on
a map.* The package exposes the same matching machinery it uses internally.

## Standardise any frame

`standardize_country()` attaches ISO codes and classifications, reconciling
spellings automatically:

```{r}
my_data <- data.frame(
  nation = c("U.S.", "S. Korea", "Czechia", "Kosovo", "Cote d'Ivoire", "UK"),
  score  = c(10, 8, 6, 4, 7, 9)
)
standardize_country(my_data, nation, warn = FALSE)
```

## One call to a map

`join_world()` auto-detects the country column, standardises it and attaches
geometry:

```{r}
my_data |>
  join_world(nation, warn = FALSE) |>
  world_map(score, title = "My data on the ISO spine")
```

## Reconcile two messy tables

`country_join()` joins two frames that each key on country names, by reconciling
both sides to `iso3c` first:

```{r}
a <- data.frame(country = c("Czechia", "South Korea", "Russia"), gdp = 1:3)
b <- data.frame(nation  = c("Czech Republic", "Korea, Rep.", "Russian Federation"),
                pop = c(10, 51, 144))
country_join(a, b, country, nation)
```

## Check before you trust

Always inspect what failed to match:

```{r}
check_country_match(my_data$nation)
```

If something legitimately cannot be matched, extend the override table:

```{r}
wdj_overrides(c(Somaliland = "SOM"))[c("Kosovo", "Somaliland")]
```

## Custom origins

If your key is already an ISO-2 or World Bank code, tell `standardize_country()`
via `origin`:

```{r}
df <- data.frame(code = c("US", "KR", "BR"))
standardize_country(df, code, origin = "iso2c", warn = FALSE)
```
